### 2024-11-24 paper/paper.qmd
fix some scripts & drew some more models

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Wendy Yuan
thanks: "Code and data are available at: https://github.com/kiwindyy/Body-Fat"
date: November 29, 2024
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(MASS)
library(GGally)
library(gridExtra)
library(knitr)
library(car)
library(kableExtra)

data <- read_parquet(here::here("data/02-analysis_data/bodyfat.parquet"))
raw_data <- read.table(here::here("data/01-raw_data/bodyfat.txt"))
```

# Introduction

Why it matters paragraph

Overview paragraph

Results paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

## Estimand

Estimand paragraph

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

Overview text

## Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

cleaned data with normalized measurement 
```{r}
#| echo: false
#| warning: false
#| message: false

# Split into two smaller tables
part1 <- data[, 1:8]  # First 8 columns
part2 <- data[, 9:16]  # Remaining 8 columns

# Render each part
kable(head(part1, 10), booktabs = TRUE, caption = "Part 1: Columns 1-8") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))

kable(head(part2, 10), booktabs = TRUE, caption = "Part 2: Columns 9-16") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

## Outcome variables

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| echo: false
#| warning: false
#| message: false

# Histogram for Pct.BF
hist(data$Pct.BF, 
     main = "Histogram of Percent Body Fat",
     xlab = "Percent Body Fat",
     col = "#f88379",
     border = "black",
     breaks = 20)
```

Talk more about it.

## Predictor variables

variables of interest: Age Height Abdomen Wrist

Reference appendix for how variables were picked using backward selection method

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

```{r}
#| echo: false
#| warning: false
#| message: false

data <- data[ , !(names(data) %in% c("Density"))]

# Full model using all variables except Density
full_model <- lm(Pct.BF ~ ., data = data)

# Perform backward selection
backward_model <- stepAIC(full_model, direction = "backward", trace = 0)

# Extract the final 4-variable model
final_variables <- names(coef(backward_model))[-1] # Exclude intercept
while (length(final_variables) > 4) {
  current_model <- lm(as.formula(paste("Pct.BF ~", paste(final_variables, collapse = "+"))), data = data)
  current_aic <- AIC(current_model)

  # Drop one variable at a time and calculate AIC
  aic_values <- sapply(final_variables, function(var) {
    temp_model <- lm(as.formula(paste("Pct.BF ~", paste(setdiff(final_variables, var), collapse = "+"))), data = data)
    return(AIC(temp_model))
  })

  # Remove the variable leading to the lowest AIC improvement
  final_variables <- setdiff(final_variables, names(which.min(aic_values)))
}

# Final 4-variable model
final_model <- lm(as.formula(paste("Pct.BF ~", paste(final_variables, collapse = "+"))), data = data)

# Print the AIC value and final variables
cat("AIC:", AIC(final_model), "\n")
cat("Variables:", final_variables, "\n")
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Create a layout for multiple plots
par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))  # 2 rows and 4 columns

# Histogram for each variable
hist(data$Age, main = "Histogram of Age", xlab = "Age", col = "#7dbeff", border = "black")
hist(data$Height, main = "Histogram of Height", xlab = "Height (inches)", col = "#2aebaa", border = "black")
hist(data$Abdomen, main = "Histogram of Abdomen", xlab = "Abdomen (cm)", col = "#eb49a4", border = "black")
hist(data$Wrist, main = "Histogram of Wrist", xlab = "Wrist (cm)", col = "#bb86f7", border = "black")
```

```{r}
#| echo: false
#| warning: false
#| message: false

model <- lm(Pct.BF ~ Age + Height + Abdomen + Wrist, data = data)

# Calculate VIF
vif_values <- vif(model)

# Print VIF values
print(vif_values)
```

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

$y_i = \beta_0 + \beta_1 x_{1i} + \beta_2 x_{2i} + \beta_3 x_{3i} + \beta_4 x_{4i}$

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

Linear reg assumptions
```{r}
#| echo: false
#| warning: false
#| message: false

# Fit the linear regression model
model <- lm(Pct.BF ~ Age + Height + Abdomen + Wrist, data = data)

# 1. Linearity: Residuals vs Fitted Plot
plot(model$fitted.values, residuals(model), 
     main = "Residuals vs Fitted",
     xlab = "Fitted Values",
     ylab = "Residuals")
abline(h = 0, col = "red")
```

```{r}
#| echo: false
#| warning: false
#| message: false

# 2. Independence: Durbin-Watson Test (requires car package)
# install.packages("car")  # Uncomment to install if not already installed
library(car)
durbinWatsonTest(model)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# 3. Homoscedasticity: Scale-Location Plot
plot(model, which = 3)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# 4. Normality: Q-Q Plot of Residuals
qqnorm(residuals(model))
qqline(residuals(model), col = "red")
```

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

Multicollinear between picked vars 
```{r}
#| echo: false
#| warning: false
#| message: false

# Select the specified variables
selected_data <- data[, c("Age", "Height", "Abdomen", "Wrist")]

# Generate a pairwise plot
pairwise_plot <- ggpairs(selected_data, 
                         lower = list(continuous = wrap("points", color = "#008080", alpha = 0.4)), 
                         upper = list(continuous = wrap("points", color = "#008080", alpha = 0.4)),
                         diag = list(continuous = wrap("densityDiag", fill = "#008080", alpha = 0.6)))

# Print the plot
print(pairwise_plot)
```

```{r}
#| echo: false
#| warning: false
#| message: false

p1 <- ggplot(data, aes(x = Age, y = Pct.BF)) +
  geom_point(color = "#7dbeff", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Pct.BF vs Age", x = "Age", y = "Pct.BF") +
  theme_minimal()

p2 <- ggplot(data, aes(x = Height, y = Pct.BF)) +
  geom_point(color = "#2aebaa", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Pct.BF vs Height", x = "Height (inches)", y = "Pct.BF") +
  theme_minimal()

p3 <- ggplot(data, aes(x = Abdomen, y = Pct.BF)) +
  geom_point(color = "#eb49a4", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Pct.BF vs Abdomen", x = "Abdomen (cm)", y = "Pct.BF") +
  theme_minimal()

p4 <- ggplot(data, aes(x = Wrist, y = Pct.BF)) +
  geom_point(color = "#bb86f7", alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Pct.BF vs Wrist", x = "Wrist (cm)", y = "Pct.BF") +
  theme_minimal()

grid.arrange(p1, p2, p3, p4, nrow = 2, ncol = 2)
```

```{r}
#| echo: false
#| warning: false
#| message: false

model <- lm(Pct.BF ~ Age + Height + Abdomen + Wrist, data = data)
summary(model)
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Fit the model
model <- lm(Pct.BF ~ Age + Height + Abdomen + Wrist, data = data)

# Extract coefficients, standard errors, and calculate 90% confidence intervals
coef_summary <- summary(model)$coefficients
conf_intervals <- confint(model, level = 0.90)  # 90% confidence intervals

# Create a data frame for plotting
coeff_data <- data.frame(
  Parameter = rownames(coef_summary),
  Estimate = coef_summary[, "Estimate"],
  Lower = conf_intervals[, 1],
  Upper = conf_intervals[, 2]
)

# Remove the intercept for better visualization (optional)
coeff_data <- coeff_data[coeff_data$Parameter != "(Intercept)", ]

# Create the plot
ggplot(coeff_data, aes(x = Estimate, y = Parameter)) +
  geom_point(size = 3, color = "#f88379") +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.2,
                 color = "#f88379", linewidth = 0.8) +
  theme_minimal() +
  labs(
    title = "Model Coefficients",
    subtitle = "90% Confidence Intervals",
    x = "Coefficient Estimate",
    y = "Predictors"
  ) +
  theme(
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold")
  )

```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

# Conclusion

\newpage

\appendix

# Appendix {-}

```{r}
#| echo: false
#| warning: false
#| message: false

# Split into two smaller tables
part1 <- raw_data[, 1:8]  # First 8 columns
part2 <- raw_data[, 9:16]  # Remaining 8 columns

# Render each part
kable(head(part1, 10), booktabs = TRUE, caption = "Part 1: Columns 1-8") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))

kable(head(part2, 10), booktabs = TRUE, caption = "Part 2: Columns 9-16") %>%
  kable_styling(latex_options = c("hold_position", "scale_down"))
```

\newpage

# References